Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-23 paper/current/main.tex
updates to conclusion

This same convex-curve repeats across many countries, though with variations that can reveal characteristics about the scientific systems of each country (Fig.~\ref{fig:length}b). 
For example, Chinese organizations have a longer average embedding vector length ($\bar{l} = 4.81$) than organizations in the United States ($\bar{l} = 0.421$), reflecting the tendency for organizations in the United States to appear more frequently in more different contexts than Chinese organizations. 
Smaller but scientifically advanced countries such as Sweden, and other large populous countries like Brazil exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset), a pattern repeated across most of the 20 countries with the most researchers (Fig.~\ref{fig:concave20}). 
Other nations exhibit different curves, such as Egypt which has an "S"-shaped curve---their largest organizations do not have smaller vectors, as in many other countries. 
Other nations such as Poland and Turkey demonstrate a linear relationship between an organization's number of researchers and the vector size, indicating that their largest organizations belong to very specific contexts (Fig.~\ref{fig:concave20}).
By encoding whether an organization will appear in specific or many different contexts, organization embeddings offer a way to characterize universal and particularistic patterns of mobility across countries.

%
% Figure - SemAxis results
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Size of organization embedding vectors captures connection diversity of organizations.}
		\textbf{a.} Size (L2 norm) of organization embedding vectors compared to their number of researchers for U.S. universities.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest ranked university.
		Uncolored points are universities not listed on a Times ranking.
		A convex-shape emerges, wherein larger universities tend to more distant from the origin (large L2 norm), however the more prestigious the university, the closer it is to the origin (small L2 norm). 
		The L2 norm is correlated weakly with the number of connection divided by the number of researchers in organization, which can be interpreted diversity of the connection compared to their size ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit, and 99\% confidence intervals are shown. 
		\textbf{b.} This convex pattern repeats globally for all organizations across all countries, shown for the United States, China, and others,
		The loess regression lines are shown for each selected country, and for the aggregate of remaining countries, with 99\% confidence intervals. 
		Loess lines are also shown for organizations in Sweden, Brazil, and Egypt (inset).
	}
	\label{fig:length}
\end{figure}

@@

Neural embeddings offer a novel approach for learning an effective and robust functional distance between places based on affiliation trajectory data that captures the complex and multi-faceted nature of mobility. 
Here, we demonstrated how the embedding space encodes aspects scientific mobility, such as the  global and regional geography, national shared languages, and the hierarchy of organizational prestige. 
While scientific mobility is used as a case study, this approach can be broadly applied to other kinds of mobility data, such as general human migration, transit-network mobility, and more.
Moreover, this approach can be used to learn a functional distance between entities for which no geographic analog exists, such as between occupational categories based on individual's career trajectories. 
In addition to providing a functional distance that supports modeling and predicting mobility patterns, the unique structure of the neural embedding space is amenable to a range of unique applications for studying mobility. 
As we have shown, projecting the embedding space onto two dimensions using UMAP~\autocite{mcinnes2018umap} allows for the visualization of the complex structure of scientific mobility at high resolution and at multiple scales, providing the largest and most detailed map of global scientific mobility to date. 
Other techniques, such as SemAxis~\autocite{an2018semaxis}, also allow the potential to quantify abstract notions in mobility such as academic prestige, and can also be generalized to other abstract axes.
Investigation of the structure of the embedding space, such as the vector length, reveals unique and universal patterns based on organization's size and connectedness, and could contribute to deeper understandings of scientific mobility. 
Neural embedding opens the door for new approaches to represent and study scientific mobility, and also mobility more generally.

>>>>

This same convex-curve repeats across many countries, though with variations that can reveal characteristics about the scientific systems of each country (Fig.~\ref{fig:length}b). 
For example, Chinese organizations have a longer average embedding vector size ($\bar{l} = 4.81$) than organizations in the United States ($\bar{l} = 0.421$), reflecting the tendency for organizations in the United States to appear more frequently in more different contexts than Chinese organizations. 
Smaller but scientifically advanced countries such as Sweden, and other large populous countries like Brazil exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset), a pattern repeated across most of the 20 countries with the most researchers (Fig.~\ref{fig:concave20}). 
Other nations exhibit different curves, such as Egypt which has an "S"-shaped curve---their largest organizations do not have smaller vectors, as in many other countries. 
Other nations such as Poland and Turkey demonstrate a linear relationship between an organization's number of researchers and the vector size, indicating that their largest organizations belong to very specific contexts (Fig.~\ref{fig:concave20}).
By encoding whether an organization will appear in specific or many different contexts, organization embeddings offer a way to characterize universal and particularistic patterns of mobility across countries.

%
% Figure - Boomerang
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Size of organization embedding vectors captures connection diversity of organizations.}
		\textbf{a.} Size (L2 norm) of organization embedding vectors compared to their number of researchers for U.S. universities.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest ranked university.
		Uncolored points are universities not listed on a Times ranking.
		A convex-shape emerges, wherein larger universities tend to more distant from the origin (large L2 norm), however the more prestigious the university, the closer it is to the origin (small L2 norm). 
		The L2 norm is correlated weakly with the ratio between organization's of number of connected organizations and its number of researchers ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit, and 99\% confidence intervals are shown. 
		\textbf{b.} This convex pattern repeats globally for all organizations across all countries, shown for the United States, China, and others,
		The loess regression lines are shown for each selected country, and for the aggregate of remaining countries, with 99\% confidence intervals. 
		Loess lines are also shown for organizations in Sweden, Brazil, and Egypt (inset).
	}
	\label{fig:length}
\end{figure}

@@

Neural embeddings offer a novel, data-driven solution for learning an effective and robust functional distance between places based on trajectory data.
By virtue of being data-driven, and by the nature of neural embedding, it encodes the complex and multi-faceted nature of mobility. 
Here, we demonstrated how the embedding space encodes various aspects of scientific mobility, such as the  global and regional geography, national shared languages, and the hierarchy of organizational prestige, even without explicit information of these factors.
While scientific mobility is used as a case study, this approach can be broadly applied to other kinds of mobility data, such as general human migration, transit-network mobility, and more.
Moreover, this approach can be used to learn a functional distance between entities for which no geographic analog exists, such as between occupational categories based on individual's career trajectories. 
In addition to providing a functional distance that supports modeling and predicting mobility patterns, the unique structure of the neural embedding space is amenable to a range of unique applications for studying mobility. 
As we have shown allows the visualization of the complex structure of scientific mobility at high resolution across multiple scales, providing the largest and most detailed map of global scientific mobility to date. 
Other techniques, such as SemAxis~\autocite{an2018semaxis}, also allow us to quantify abstract notions in mobility such as academic prestige, and can also be generalized to other abstract axes.
Investigation of the structure of the embedding space, such as the vector size, reveals unique and universal patterns based on organization's size and connectedness, and could contribute to deeper understandings of scientific mobility. 
Neural embedding opens the door for new approaches to represent and study scientific mobility, and also mobility more generally.
